Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Swept through in reverse, added future work, some related work

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{fig:control} and discussion).
This reflects a cleaner division between data protection and control-flow protection
than offered by LSE and WBCF.

@@

Unless otherwise noted\bcp{where is it ``otherwise noted''??}, colored boxes
indicate an assignment of the address to a
{\em domain} -- a label reflecting a particular set of access rights. Domains
are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
variables. Domains are part of the {\em context} of a state.
We will introduce the remaining context as needed.

@@

\begin{subfigure}{.2\textwidth}
  {\small
\begin{verbatim}
int main() {
  int x;
  x = f();
  return g();
}

@@

int g() {
  int z;
  if (z == 5)
    return 10;
  else
    return 20;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.7\textwidth}
  {\small
\begin{verbatim}
1 main:  add $1,%sp      ; allocate frame
2        call f
3        mov %ra,-1(%sp) ; store x
4        call g
5        sub $1,%sp      ; deallocate frame
6        ret

@@

20 g:    add $1,%sp      ; allocate frame
21       cmp $5,-1(%sp)  ; compare
22       brneq #25       ; branch if not equal
23       mov $10,%ra     ; set return value
24       jmp #26
25       mov $20,%ra     ; set return value
26       sub $1,%sp      ; deallocate frame
27       ret
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 22} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
  \\ &
  {\bf 23}
  ~$\cdots$
  &
  {\bf 25}
  ~$\cdots$
  \\
\end{tabular}
\end{center}

@@

\caption{Another confidentiality violation \bcp{Again, let's align the C
    code with the assembly, etc.}}
\label{fig:conf2}
\end{figure}

@@

@@

The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
in this case one that holds on any state whose program counter points to the correct instruction.
We introduce a new piece of context: we keep a stack of return targets, and a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, here represented by sharing a color with the sealed data.

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
%
\bcp{(Moved this here, BTW.)}Recall that we keep track of which addresses are
annotated as calls, with a ``code map'' $\codemap \in \CODEMAPS \subseteq
\WORDS$.
\apt{Doesn't match previous definition any more.}

\judgmentthree[ Call]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
              {\(\components = \sealcon ~ \mach\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

@@

\definition A system enjoys {\em stepwise stack integrity} if, for any reachable state
\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
\pi_\mach(\mpcstatename) ~ \component\).

@@

For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
let \(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \outside\).
A system enjoys \emph{stepwise stack confidentiality} if,
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
\(\MPCT \doteq \NPCT\).

@@

\apt{This is way better up until the final definition, which still needs more unpacking in words.}
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

@@

A system enjoys {\em return-to-caller} when, for every pair of reachable states
\(\mpcstatename = (\_,\_,(\_,\rts))\) and
\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), \(|\rts| - 1 \leq |\rts'| \leq |\rts| + 1\).

@@

Next we introduce variable passing. We will need to add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed -- when passing by value, this will typically give the memory cells
at the appropriate offset from the stack pointer, while passing by reference gives the
addresses of the appropriate variables wherever they happen to be.\apt{Why can't
  these also be given as offsets from stack pointer (as in the example)?}
\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
And we extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
\times (\rts \in \listT (\target \subseteq \MACHS)) \]
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map when pushing it, where \(\depth\) is an integer depth.
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
                                 \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

\paragraph*{Deferred Stepwise Integrity}

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).

Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\) -- meaning that the
call does return -- let \(\components'\) be the set of components that differ
between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
\(\nach \hookrightarrow \NPCT\). A system enjoys
{\em deferred stepwise integrity} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).

@@

We now can sketch observational versions of both integrity and confidentiality.
Integrity is already straightforward, simply weakening the condition that must hold
after the return.

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component = \sealed{\_}\).

Then, again supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components'\) be the set of components that differ
between \(\mach\) and \(\mach'\). Take any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
\(\nach \hookrightarrow \NPCT\). A system enjoys
{\em observational integrity} if for all such \(\MPCT\) and \(\NPCT\),
\(\obsof(\MPCT) \simeq \obsof(\NPCT)\).

We amend our intuitive statement to: ``No visible reads after dangerous writes.''

@@

\definition Let \(\components\) be a set of components and \(\mach,\mach'\) and \(\nach,\nach'\)
be pairs of states. The {\em unsafe set} of \(\components\), written
\(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
\(\component \in \components\) such that if
\(\mach ~ \component \not = \mach' ~ \component\) or
\(\nach ~ \component \not = \nach' ~ \component\), then
\(\mach' ~ \component = \nach' ~ \component\).

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Let \(\nach\) be a variant such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_\depth \).

A system enjoys {\em observational confidentiality} if for all such
\(\MPCT\) and \(\NPCT\), the following conditions all hold:

@@

\sna{More to write here. The gist is: this is the extensional property.
  The one that we would tentatively propose as a real characterization of stack
  safety.}

@@

    Observable properties allow us to defer enforcement until a property
    violation would become visible, as in lazy policies. But by formalizing
    the connection to the eager policy, we can identify when lazy policies
    miss violations that can become visible. The optimized lazy tag policy from
    \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
    at which it was written, and enforces that it must be read from the
    same depth. However, this allows violations of observable stack safety!
    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
    write and use it to create an observable difference that violates both
    observable integrity and confidentiality without violating the policy.

    The lazy policy can be repaired if, instead of tagging each function
    activation with its depth in the stack, we generate a fresh activation
    identifier on each call, which prevents a program from exploiting stale
    tags from previous activations. A related mechanism was explored in the
    Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
    associates a unique activation identifier to each function, which is
    however shared by all activations of the same function in order to obtain
    good cacheability of tag rules. Static Authorities comes closer to
    observable stack safety, though the example above would still exist
    if {\tt bar} and {\tt baz} were merged into a single function that tracked
    whether it had been called previously and changed its behavior accordingly.
    \ifaftersubmission\apt{More blatently, static authorities doesn't work for recursive calls.}\fi

@@

StkTokens \citep{Skorstengaard+19} continues this line of work. Like
the earlier paper, it aims to protect the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a new
calling convention that makes use of linear capabilities for stack and return
pointers. The convention operates on a single shared stack and requires that
protected components avoid compromising their own security by following certain
simple rules---like not leaking their private capability seals. More
precisely but still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces; their model of
observations is limited to cotermination.

We conjecture that the linear capability machines introduced by
\citep{Skorstengaard+19}, in combination with the StkTokens calling convention,
can be modeled in our framework and satisfy our definition of stack safety in
\cref{sec:lse}. Further, we expect them to satisfy the even stronger
property developed in \cref{sec:enforcement}. Proving (or testing) these conjectures would
involve exposing their formal definitions of local state encapsulation and
well-bracketed control flow, which are built-in as part of the
semantics of the capability machines and not given explicitly.
The main practical limitation of StkTokens
is its reliance on linear capabilities, as it is unclear now they could be added
to practical capability machines, especially in terms of efficiency (previous
work on micro-policies \citep{yannis-report} has shown how to use that framework
to implement linear return capabilities.\bcp{But IIUC the Dover realization
  of micro-policies can't do this...?})\rb{Correct}
%
In recent work, \citet{Georges+21} address these efficiency concerns
with the introduction of a new type of uninitialized capability that
avoids the need to clear the stack while being more amenable to
practical hardware implementation, although like
\citet{Skorstengaard+19b} it lacks a general stack safety theorem,
instead using a similar logical relation to reason about the safety of
individual programs.

@@

@@

  We have presented a pure notion of stack safety that omits many complicating
  factors common to real systems. In particular, our model as presented exploits
  the simplifying assumption that privilege never increases during a function
  call, and that therefore a contour computed once on the entry to a function
  describes its privilege throughout. Our future plans are to extend the model
  with common language features, some of which violate this assumption. Here we
  describe the basic principles by which our model can be extended to some of
  these cases. We also plan to develop a testing framework for quickly
  checking the validity of different enforcement mechanisms.

>>>>

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{sec:control}).
This reflects a cleaner division between data protection and control-flow protection
than offered by LSE and WBCF.

@@

Colored boxes indicate an assignment of the address to a
{\em domain} -- a label reflecting a particular set of access rights. Domains
are \(\unsealed\)\bcp{``unsealed'' would read better} (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
variables. Domains are part of the {\em context} of a state.
We will introduce the remaining context as needed.

@@

\begin{subfigure}[t]{.2\textwidth}
  {\small
\begin{verbatim}
int main() {
  int x;
  x = f();
  return g();
}

@@

int g() {
  int z;
  return z;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.7\textwidth}
  {\small
\begin{verbatim}
1 main:  add $1,%sp      ; allocate frame
2        call f
3        mov %ra,-1(%sp) ; store x
4        call g
5        sub $1,%sp      ; deallocate frame
6        ret

@@

20 g:    add $1,%sp      ; allocate frame
21       mov -1(%sp),%ra ; violation!
22       sub $1,%sp      ; deallocate frame
23       ret
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{62}
  \MemoryLabel{-10em}{0.75em}{105}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 22} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{\bf 5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{36}
  \MemoryLabel{-10em}{0.75em}{\bf 8}
  \MemoryLabel{-6em}{0.75em}{90}
  \\
\end{tabular}
\end{center}

@@

\caption{Another confidentiality violation}
\label{fig:conf2}
\end{figure}

@@

\label{sec:controlflow}

@@

The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate return target in {\tt main}.
We add a new piece of context to our diagram: we keep a stack of return targets, and a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, here represented by sharing a color with the sealed data.

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a predicate on addresses.

\judgmentthree[ Call]
              {\(\codemap ~ (\mach ~ \PCname)\)}
              {\(\components = \sealcon ~ \mach\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
                {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

@@

\definition {\bf Stepwise Stack Integrity}

A system enjoys {\em stepwise stack integrity} if, for any reachable state
\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and any component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename') ~ \component =
\pi_\mach(\mpcstatename) ~ \component\).

@@

For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
let \(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \outside\). That is, at the start of some call, we take every
component in the stack.
A system enjoys \emph{stepwise stack confidentiality} if,
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
\(\MPCT \doteq \NPCT\) -- if every variant (with the same policy state)
steps in lockstep until both traces have reached a point whose depth is
below the starting one. This is a return.

@@

Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

@@

A system enjoys {\em return-to-caller} when, for every pair of reachable states
\(\mpcstatename = (\_,\_,(\_,\rts))\) and
\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|\)

@@

Next we introduce variable passing. We will need to add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed -- when passing by value, this will typically be freshly allocated
cells at the top of the frame, while passing by reference just gives the
addresses of the appropriate variables.
\[\codemap \in \CODEMAPS ::= \addr \in \WORDS \rightharpoonup \callmap(P \subseteq \MACHS \times \COMPONENTS)\]
And we extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
\times (\rts \in \listT (\target \subseteq \MACHS)) \]
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map when pushing it, where \(\depth\) is an integer depth.
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
                                 \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

\paragraph*{Deferring Integrity Checks}

@@

\definition {\bf Deferred Stack Integrity}

For any \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \sealed{\_}\).
Then if \(\last(\MPCT) = \mach',\pol',\context'\), let \(\components' = \Delta(\mach,\mach')\).
 A system enjoys {\em deferred stepwise integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), \(\mach \hookrightarrow \MPCT\), and
\(\nach \hookrightarrow \NPCT\), \(\MPCT \doteq \NPCT\).

@@

We now can sketch observational versions of both integrity and confidentiality.
Integrity is already straightforward, simply weakening the condition that must hold
after the return -- instead of preventing dangerous writes from being read, we
prevent them from being leaked.

\definition {\bf Observational Stack Integrity}

For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\)
such that \(\domMap ~ \component = \sealed{\_}\).

Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components'\) be the set of components that differ
between \(\mach\) and \(\mach'\). A system enjoys
{\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), \((\mach,\pol',\context') \hookrightarrow \MPCT\), and
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT) \simeq \obsof(\NPCT)\).

@@

\definition Let \(\components\) be a set of components and \(\mach,\mach'\) and \(\nach,\nach'\)
be pairs of states. The {\em unsafe set} of \(\components\), written
\(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
\(\mach' ~ \component \not = \nach' ~ \component\).

\definition {\bf Observational Stack Confidentiality}

For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\)
such that \(\domMap ~ \component \not = \outside\).

A system enjoys {\em observational confidentiality} if for all
\(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_\depth \),
\(\MPCT\) and \(\NPCT\), the following conditions all hold:

@@

Together with return-after-call, these two properties give us
{\em Observable Stack Safety}. This is a weaker property
than the stepwise one, and also much more extensional.
We can think of stepwise, deferred stepwise, and observational as
three points along a spectrum from eager to lazy. The first two
are useful for understanding particular enforcement mechanisms,
but if we had to choose a single property to represent stack safety,
it would be observational state safety.

@@

Observable properties allow us to defer enforcement until a property
violation would become visible, as in lazy policies. But by formalizing
the connection to the eager policy, we can identify when lazy policies
miss violations that can become visible. The Lazy Tagging and Clearing
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity.

Lazy Tagging and Clearing can be repaired if, instead of tagging each function
activation with its depth in the stack, we generate a fresh activation
identifier on each call, which prevents a program from exploiting stale
tags from previous activations. A related mechanism was explored in the
Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
associates a unique activation identifier to each function, which is
however shared by all activations of the same function in order to obtain
good cacheability of tag rules, but that policy allows leaks between
instances of the same function.

Lazy Per-activation Tagging and Clearing ends up enforcing the combination
of deferred stepwise stack integrity and stepwise confidentiality -- we
don't actually need a deferred confidentiality property here. Therefore,
it also enforces observational stack safety, so we declare it stack safe.
Our properties have allowed us to reject an enforcement mechanism as
not really stack safe, and describe a fix.

@@

StkTokens \citep{Skorstengaard+19} continues this line of work. Like
the earlier paper, it aims to protect the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a new
calling convention that makes use of linear capabilities for stack and return
pointers. The convention operates on a single shared stack and requires that
protected components avoid compromising their own security by following certain
simple rules---like not leaking their private capability seals. More
precisely but still informally, local state encapsulation is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{Skorstengaard+19b} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces; their model of
observations is limited to cotermination.
\sna{Some of this seems like more detail than we need, and we need to
  compare ourselves to their list of criteria.}

@@

We have evaluated Depth Isolation, Lazy Tagging, and Lazy Clearing against
our properties, finding that Depth Isolation does indeed enforce stepwise
stack safety and therefore also observable stack safety. We also found that
neither lazy micro-policy actually does enforce stack safety, and proposed
a fix.

@@

We have introduced a characterization against which stack safety enforcement
mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
From here there are two natural ways to extend the work. We should thoroughly evaluate
other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
the functionality of our definitions by showing that they can be used to prove
the correctness of a programming logic with strengthened rules around calls and returns.

\paragraph*{Stack Safety in the Cheri Zoo}

There are several proposals around the use of Cheri capabilities to enforce stack safety,
ranging from mechanisms that use the standard Cheri hardware \citep{Skorstengaard+19b},
to those that propose entirely new types of capabilities, such as linear
(\citep{Skorstengaard+19}), uninitialized (\citep{Georges+21}), and monotonic (\sna{cite}).
Of these, uninitialized capabilities seem particularly promising -- they naturally enforce
write-before-read behavior, and therefore will enable confidentiality without the need
to clear large swaths of the stack. We propose to implement a model of the stack safety
system described in (\citep{Georges+21}) and test it against our properties using
QuickChick.
